---
title: "The Modeltime Spark Backend"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The Modeltime Spark Backend}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  
  out.width='100%',
  fig.align = "center",
  fig.width = 7,
  fig.height = 5,
  
  message = FALSE,
  warning = FALSE
)
```

__Modeltime is built for scale__, and as such many users may want to use a distributed backend that allows for scale beyond their local machine. One such tool is [__Spark__](https://spark.apache.org/), a unified engine for large-scale data analytics. In this tutorial, we'll show you how to use the __Modeltime Spark Backend__ for high-performance computing. 

# Spark for Iterative (Nested) Forecasting

One of the most common situations that parallel computation is required is when doing __iterative forecasting__ where the data scientist needs to experiment with 10+ models across 10,000+ time series. It's common for this large-scale, high-performance forecasting exercise to take days. 

We'll show you how we can __combine Modeltime Nested Forecasting and it's Parallel Spark Backend__ to scale this computation with distributed parallel Spark execution. Let's go!

# System Requirements

This tutorial requires:

- __sparklyr__: Used to register the spark `foreach` adapter via `registerDoSpark()`.

- __Java__: Spark installation depends on Java being installed. 

- __Spark Installation:__ Can be accomplished via `sparklyr::spark_install()` provided the user has `sparklyr` and Java. 

# Libraries

Load the following libraries. 

```{r}
library(sparklyr)
library(modeltime)
library(timetk)
library(tidymodels)
library(dplyr)
```

# Spark Connection

Next, we set up a Spark connection via `sparklyr`. For this tutorial, we use the "local" connection. But many users will use [Databricks](https://databricks.com/) to scale the forecasting workload. 

To run Spark locally:

```{r, include = FALSE}
# Run the Spark chunks only when a local Spark connection can be made.
# Otherwise the code is shown without being run, so the article still builds.
sc <- tryCatch(
  spark_connect(master = "local"),
  error = function(e) NULL
)
knitr::opts_chunk$set(eval = !is.null(sc))
```

```{r, eval = FALSE}
sc <- spark_connect(master = "local")
```

If using Databricks, you can use:

```{r, eval = F}
sc <- spark_connect(method = "databricks")
```

# Setup the Spark Backend

Next, we register the Spark Backend using `parallel_start(sc, .method = "spark")`. This is a helper to set up the `registerDoSpark()` foreach adaptor. In layman's terms, this just means that we can now run  parallel using Spark. 

```{r}
parallel_start(sc, .method = "spark")
```

# Data Preparation (Nested Forecasting)

We'll run through a minimal nested forecasting example. Refer to the [Nested Forecasting Tutorial](nested-forecasting.html) for further information on how to perform iterative forecasting with the "nested" data structure in `modeltime`.

The dataset we'll be forecasting is the `walmart_sales_weekly`, which we modify to just include 3 columns: "id", "date", "value". 

1. The __id__ feature is the grouping variable. 
2. The __date__ feature contains timestamps.
3. The __value__ feature is the sales value for the Walmart store-department combination. 

```{r}
walmart_sales_weekly %>%
    dplyr::select(id, date = Date, value = Weekly_Sales) %>%
    dplyr::group_by(id) %>%
    plot_time_series(date, value, .facet_ncol = 2, .interactive = FALSE)
```

We prepare as nested data using the Nested Forecasting preparation functions. 

```{r}
nested_data_tbl <- walmart_sales_weekly %>%
    dplyr::select(id, date = Date, value = Weekly_Sales) %>%
    extend_timeseries(
        .id_var        = id,
        .date_var      = date,
        .length_future = 52
    ) %>%
    nest_timeseries(
        .id_var        = id,
        .length_future = 52
    ) %>%
    
    split_nested_timeseries(
        .length_test = 52
    )

nested_data_tbl
```

# Modeling

We'll create two unfitted models: XGBoost and Prophet. Then we'll use `modeltime_nested_fit()` to iteratively fit the models to each of the time series using the Spark Backend. 

## XGBoost

We create the XGBoost model on features derived from the date column. 

```{r}
rec_xgb <- recipe(value ~ ., extract_nested_train_split(nested_data_tbl)) %>%
    step_timeseries_signature(date) %>%
    step_rm(date) %>%
    step_zv(all_predictors()) %>%
    step_dummy(all_nominal_predictors(), one_hot = TRUE)

wflw_xgb <- workflow() %>%
    add_model(boost_tree("regression") %>% set_engine("xgboost")) %>%
    add_recipe(rec_xgb)

wflw_xgb
```
## Prophet

Next, we create a prophet workflow. 

```{r}
rec_prophet <- recipe(value ~ date, extract_nested_train_split(nested_data_tbl)) 

wflw_prophet <- workflow() %>%
    add_model(
        prophet_reg("regression", seasonality_yearly = TRUE) %>% 
            set_engine("prophet")
    ) %>%
    add_recipe(rec_prophet)

wflw_prophet
```

# Nested Forecasting with Spark

Now, the beauty is that everything is set up for us to perform the nested forecasting with Spark. We simply use `modeltime_nested_fit()` and make sure it uses the Spark Backend by setting `control_nested_fit(allow_par = TRUE)`.

Note that this will take about 30-seconds because we have a one-time cost to move data, libraries, and environment variables to the Spark clusters. But the good news is that when we scale up to 10,000+ time series, that the one-time cost is minimal compared to the speed up from distributed computation. 

```{r, message=TRUE}
nested_modeltime_tbl <- nested_data_tbl %>%
    modeltime_nested_fit(
        wflw_xgb,
        wflw_prophet,
        
        control = control_nested_fit(allow_par = TRUE, verbose = TRUE)
    )
```

The nested modeltime object has now fit the models using Spark. 

```{r}
nested_modeltime_tbl
```

## Model Test Accuracy

We can observe the results. First, we can check the accuracy for each model

```{r}
nested_modeltime_tbl %>%
  extract_nested_test_accuracy() %>%
  table_modeltime_accuracy(.interactive = F)
```

## Test Forecast

Next, we can examine the test forecast for each of the models. 

```{r}
nested_modeltime_tbl %>%
  extract_nested_test_forecast() %>%
  dplyr::group_by(id) %>%
  plot_modeltime_forecast(.facet_ncol = 2, .interactive = F)
```

# Close Clusters and Shutdown Spark

We can close the Spark adapter and shut down the Spark session when we are finished. 

```{r}
# Unregisters the Spark Backend
parallel_stop()

# Disconnects Spark
spark_disconnect_all()
```


# Summary

This article ran a nested forecast using Spark as the parallel processing backend.
